Viscous Erosion in Porous Media

### 2018-03-25 docs/methods/erosion_methods.tex
small edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shear stress}
\label{sec:shearStress}
\todo[inline]{Bryan is here on editing}
We have decomposed the deformation tensor into contributions from three
different terms.  The contributions from the Stokeslets and the rotlets
can be directly evaluated since $\rr \neq 0$ for all $\xx \in \gamma$.
For the shear stress of the double-layer potential, we would like to use
the trapezoid rule. The kernel of the deformation tensor due to
the double-layer potential~\eqref{eqn:shearStressDLP} is 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right).
\end{align*}
Unfortunately, this kernel has a singularity that is asymptotic to
$\bigO(\rho^{-2})$ when $\yy \rightarrow \xx$, and this is too strong
for the trapezoid rule.  The singularity strength can be reduced since
the velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression for the deformation tensor has a singularity that is
asymptotic to $\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Finally,
to achieve spectral accuracy, we apply odd-even
integration~\cite{sid-isr1988}.  Therefore, the deformation tensor of
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_\ell$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_\ell}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_\ell \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  The deformation
tensor due to the Stokeslets and rotlets are given in
equation~\eqref{eqn:shearStressSR}.  Finally, the jump in the
deformation tensor~\eqref{eqn:shearStressJump} is computed with Fourier
differentiation.  Finally, the shear stress is computed by multiplying the
deformation tensor by the normal vector, and taking the inner product
with the tangent vector.

We do not employ an FMM to evaluate the shear stress at $\xx_i$,
$i=1,\ldots,N$, so the cost of evaluating the shear stress is
$\bigO(N^2)$.  However, this calculation is only performed once per time
step as opposed to the velocity double-layer potential that must be
evaluated at each GMRES iteration.  In future work, we plan to use a
kernel-independent FMM~\cite{yin-bir-zor2004} to evaluate the shear
stress with $\bigO(N)$ operations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Post processing quantities of interest}
\label{s:qoi}
To study characteristics of the flow, the vorticity
(Section~\ref{sec:vorticity}), pressure (Section~\ref{sec:pressure}),
and drag (Section~\ref{sec:drag}) will be computed.  Once
equation~\eqref{eqn:completed_DLP} is solved for the density function,
Stokeslets, and rotlets, all of these quantities can be computed.  For
target points in the fluid domain, the resulting layer potentials
involve integrands that are periodic and smooth, so the trapezoid rule
guarantees spectral accuracy~\cite{tre-wei2014}.  However, if the target
point is close to $\bd \Omega$, a nearly-singular integrand must be
integrated, and this requires upsampling that is described in
Section~\ref{sec:NSI}.  When the target points is on the boundary, the
integrands involve singularities that are evaluated by combining
singularity subtraction and odd-even integration~\cite{sid-isr1988}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Shear stress}
\label{sec:shearStress}
We have decomposed the deformation tensor into contributions from three
different terms.  The contributions from the Stokeslets and the
rotlets~\eqref{eqn:shearStressSR} can be immediately evaluated since
they do not involve any integration that needs to be approximated.  For
the shear stress of the double-layer potential, we use a slight variant
of the trapezoid rule. The kernel of the deformation tensor due to the
double-layer
potential~\eqref{eqn:shearStressDLP} is 
\begin{align*}
  K_\ssigma(\xx,\yy)\eeta(\yy) &= \frac{1}{2\pi} \left(
    2\frac{\rr \cdot \nn}{\rho^2} \frac{\rr \cdot \eeta}{\rho^2} \II + 
    \frac{\rr \cdot \eeta}{\rho^4} (\nn \otimes \rr + \rr \otimes \nn)
    \right. \\ & \hspace{25pt}+ \left.
    \frac{\rr \cdot \nn}{\rho^4} (\eeta \otimes \rr + \rr \otimes \eeta) - 
    8\frac{(\rr \cdot \nn)(\rr \cdot \eeta)}{\rho^6}(\rr \otimes \rr)
  \right).
\end{align*}
This kernel has a singularity that is asymptotic to $\bigO(\rho^{-2})$
when $\yy \rightarrow \xx$, and this is too strong for the unmodified
trapezoid rule.  The singularity strength can be reduced since the
velocity corresponding to a constant density function is
constant~\cite{poz1992}.  Therefore, the deformation tensor of the
double-layer potential is equivalent to
\begin{align*}
  \ssigma^{\DD}(\xx) = \int_{\bd\Omega}K_\ssigma(\xx,\yy)
      (\eeta(\yy) - \eeta(\xx)) \, ds_{\yy}, \quad \xx \in \bd\Omega.
\end{align*}
This expression has a singularity that is asymptotic to
$\bigO(\rho^{-1})$ when $\yy \rightarrow \xx$.  Then, to achieve
spectral accuracy, we apply the trapezoid rule with odd-even
integration~\cite{sid-isr1988}.  Therefore, the deformation tensor of
the double-layer potential at a discretization point $\xx_i$ on body
$\gamma_\ell$, with $i$ being odd, is approximated as
\begin{align}
  \ssigma^{\DD}(\xx_i) = \sum_{\xx_j \notin \gamma_\ell}
    K_\ssigma(\xx_i,\xx_j) \eeta(\xx_j) \Delta s_j + 
  \sum_{\substack{\xx_j \in \gamma_\ell \\ j \: \mathtt{even}}}
    K_\ssigma(\xx_i,\xx_j) (\eeta(\xx_j) - \eeta(\xx_i)) \Delta s_j,
  \label{eqn:stressOddEven}
\end{align}
and a similar approximation is used when $i$ is even.  We also need to
approximate the jump in the deformation
tensor~\eqref{eqn:shearStressJump}, and this is done with Fourier
differentiation.  Finally, the shear stress is computed by multiplying
the deformation tensor by the normal vector, and taking the inner
product with the tangent vector.

We do not employ an FMM to evaluate the shear stress at $\xx_i$,
$i=1,\ldots,N$, so the cost of evaluating the shear stress is
$\bigO(N^2)$.  However, this calculation is only performed once per time
step as opposed to the velocity double-layer potential, where we do
employ the FMM, that must be evaluated at each GMRES iteration.  In
future work, we plan to use a kernel-independent
FMM~\cite{yin-bir-zor2004} to evaluate the shear stress with $\bigO(N)$
operations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Post processing quantities of interest}
\label{s:qoi}
\todo[inline]{Bryan is here}
To study characteristics of the flow, the vorticity
(Section~\ref{sec:vorticity}), pressure (Section~\ref{sec:pressure}),
and drag (Section~\ref{sec:drag}) will be computed.  Once
equation~\eqref{eqn:completed_DLP} is solved for the density function,
Stokeslets, and rotlets, all of these quantities can be computed.  For
target points in the fluid domain, the resulting layer potentials
involve integrands that are periodic and smooth, so the trapezoid rule
guarantees spectral accuracy~\cite{tre-wei2014}.  However, if the target
point is close to $\bd \Omega$, a nearly-singular integrand must be
integrated, and this requires upsampling that is described in
Section~\ref{sec:NSI}.  When the target points is on the boundary, the
integrands involve singularities that are evaluated by combining
singularity subtraction and odd-even integration~\cite{sid-isr1988}.
